---
title: "SLiMSuite REST API"
author: "Rich Edwards"
date: "18/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Programmatic access to SLiMSuite REST services

By default, the REST services return human-readable HTML when queued or running. Default output is formatted full server output.

Individual components can be accessed in plain text using &rest=X, where X is the chosen server output. Details of server output can be found in the main Help page for universal outputs, or by running the chosen program with &rest=outfmt. REST outputs are also listed in the main documentation for the program, accessed with &rest=docs.

However, even when plain text output is being required, the main program run may still need to go into a queue and/or may take several minutes to complete. To avoid proxy timeouts, these outputs will still go via a human-readable intermediate queue page.

Programmatic access is therefore a three step process:

1. Set the job going and retrieve the `JobID` by calling the program with `&rest=jobid`. This will return the job ID in the form of an 11-digit number, e.g. `15010900002`.

2. Check the progress of your job using `check&jobid=X`, where `X` is the job ID returned in step 1. This will return `Queued`, `Running` or `Finished`, dependent on the status of the job. If something has gone wrong, `Failed` will be returned. Please cycle and check periodically until `Finished` is return. We request that you do not make this cycle period too short - once every few seconds/minutes (depending on the job) is recommended.

3. Once `Finished`, retrieve the desired output using `retrieve&jobid=X&rest=X`. To get all outputs, use `&rest=full`. The tool `SLiMParser` can be used to split the full output into separate files if desired. If there is a problem, the output should start with `ERROR:`.

Note: By default, jobs are unprotected. To eliminate the risk of other users retrieving your jobs, you can add `&password=X` to each of these calls.

SLiMParser can also be used to execute this process in full:

```
python SLIMSUITEPATH/tools/slimparser.py restin=RESTURL restout=T restoutdir=PATH
```

## Example R Code for Programmatic Access

We will be using the mock ["Zen Wisdom Generator"](http://rest.slimsuite.unsw.edu.au/zen) as an example as it is quick to run. (**Disclaimer:** `rje_zen` was written to generate entertaining errors during testing and word combinations are random - it is not to be taken seriously in _any_ way!)

First, we need to set up the URL for the server call itself (see <http://rest.slimsuite.unsw.edu.au>). The basic REST commands all have the same structure:

```
http://rest.slimsuite.unsw.edu.au/prog&options
```

where `prog` is the SLiMSuite program to run and `&options` is a list of options separated by `&` characters in the form `&cmd=value`. Options match those documented in the relevant SLiMSuite program, e.g. the command line option `dismask=T` would be given as `&dismask=T`. In addition, the server can take `&rest=X` to specificy the output and/or `&password=X` to protect the job with a password.

**Note:** Without at least one `&` in the URL, it be recognised as a documentation request, e.g. (`http://rest.slimsuite.unsw.edu.au/zen`)[http://rest.slimsuite.unsw.edu.au/zen].

```{r rest}
restbase = "http://rest.slimsuite.unsw.edu.au/"
restprog = "zen"
# We will set the server to return 10 wisdoms and sleep for 1s between wisdoms, so the job should take about 10 seconds.
restopt = "&wisdoms=10&zensleep=1"
#i# Full REST call:
restcall = paste0(restbase,restprog,restopt)
writeLines(restcall)
```

Next, we need to call the server and retrieve the job ID:

```{r jobid}
jobcall = paste0(restcall,"&rest=jobid")
writeLines(jobcall)
jobid = readLines(jobcall,warn=FALSE)
writeLines(jobid)
```

Then, we check the status of the job. Jobs can be `Queued`, `Running`, `Finished` or `Failed`, so we will want to loop until it is complete:

```{r checkstatus}
jobcall = paste0(restcall,"&rest=jobid")
writeLines(jobcall)
jobid = readLines(jobcall,warn=FALSE)
joburl = paste0(restbase,"check","&jobid=",jobid)
writeLines(joburl)
status = readLines(joburl,warn=FALSE)
while(! status %in% c("Finished","Failed")){
  writeLines(paste0(status," - Sleep!"))
  Sys.sleep(5)   # We will pause for a five seconds to give the job a chance
  status = readLines(joburl,warn=FALSE)
}
writeLines(status)
```
**Note:** Bigger jobs will be queued for a while before Running. It is not currently possible to predict the length of time that jobs might be queued. You can see the current queue status at <http://rest.slimsuite.unsw.edu.au/status>.

Once the job is run, you will want to retrieve the results. A list of possible outputs for each server can be retrieved with `&rest=restkeys`:

```{r restkeys}
joburl = paste0(restbase,"retrieve&jobid=",jobid,"&rest=restkeys")
writeLines(joburl)
restkeys = readLines(joburl,warn=FALSE)
print(restkeys)
```
The following are generic for all servers: `status`, `version`, `ini`, `log`, `warnings`, `errors`, `prog`. Not all of these appear in the `restkeys` list.

Additional outputs are server-specific. In this case, we have `wisdoms`. These can be read simply into a character vector like the `restkeys` output itself, using `readLines()`:

```{r wisdoms}
restkey = "wisdoms"
joburl = paste0(restbase,"retrieve&jobid=",jobid,"&rest=",restkey)
writeLines(joburl)
wisdoms = readLines(joburl,warn=FALSE)
writeLines(wisdoms)
```
It is also possible to read outputs into a data frame using `read.delim()`. The `log` output, for example, is a tab-delimited file (without headers):

```{r logtable}
restkey = "log"
joburl = paste0(restbase,"retrieve&jobid=",jobid,"&rest=",restkey)
writeLines(joburl)
logtable = read.delim(joburl,header=FALSE,sep="\t")
colnames(logtable) = c("Log","Time","Description")
summary(logtable)
knitr::kable(logtable, caption="Tabulated log output")
```

Other servers have tabular outputs that include headers and/or are comma separated (denoted by `*.csv` extensions). The will need to be parsed with `headers=TRUE` and/or `sep=","` (or `read.csv()`).

### SLiMProb example
We can put this altogether to use a different server with an example from the [BioInfoSummer2015 SLiMSuite Workshop](http://slimsuite.blogspot.com.au/2015/12/bioinfosummer2015-slimsuite-workshop.html). This time, we will call [SliMProb](http://rest.slimsuite.unsw.edu.au/slimprob) to find a couple of custom motifs in an Uniprot protein. We will also update the code to be able to handle a password:

```{r fullapi}
#1# Setup REST Options
restbase = "http://rest.slimsuite.unsw.edu.au/"
restprog = "slimprob"
restopt = "&uniprotid=E1A_ADE02&motifs=LxCxE,PxDLS"
password = "None"   
#i# Running with None will set no password. If a password is set, it will be needed for all retrieve calls.
#># Get JobID:
restcall = paste0(restbase,restprog,restopt,"&password=",password,"&rest=jobid")
jobid = readLines(restcall,warn=FALSE)
writeLines(paste(restcall,"=",jobid))
#># Check for job completion
cycles = 0
joburl = paste0(restbase,"check","&jobid=",jobid,"&password=",password)
status = readLines(joburl,warn=FALSE)
while(! status %in% c("Finished","Failed")){
  cycles = cycles + 1
  Sys.sleep(5)   # We will pause for a five seconds to give the job a chance
  status = readLines(joburl,warn=FALSE)
}
writeLines(paste(status,"after",cycles,"sleep cycles"))
#># Print restkeys for SLiMProb server
joburl = paste0(restbase,"retrieve&jobid=",jobid,"&rest=restkeys","&password=",password)
restkeys = readLines(joburl,warn=FALSE)
print(restkeys)
writeLines( readLines(paste0(restbase,"retrieve&jobid=",jobid,"&rest=warnings","&password=",password), warn=FALSE) )
writeLines( readLines(paste0(restbase,"retrieve&jobid=",jobid,"&rest=errors","&password=",password), warn=FALSE) )
#># Retrieve SLiMProb Occ
restkey = "occ"
joburl = paste0(restbase,"retrieve&jobid=",jobid,"&rest=",restkey,"&password=",password)
occdb = read.csv(joburl)
knitr::kable(occdb, caption="SLiMProb Occurrences")
```

---



